Data preparation

### 2023-10-17 procesamiento/prod_prep.Rmd
AJUSTES 17 OCTUBRE

relig <- haven::read_dta(here::here("input/data/original/ZA7570_v2-1-0.dta"))
religTR <- relig %>% filter(country==792)

df1 <- df1 %>% rename(topbot=TOPBOT)
dfES <- dfES %>% rename(topbot=TOPBOT)
#fix problem with variable names
dfTR <- dfTR %>% rename(SEX=sex,
                        DEGREE=degree,
                        EDUCYRS=educyrs,
                        AGE=age,
                        MAINSTAT=mainstat,
                        UNION=union,
                        WEIGHT=weight,
                        URBRURAL=urbrural,
                        WORK=work,
                        HOMPOP=hompop)  
dfES$c_alphan <- NULL
dfTR$c_alphan <- NULL
df2 <- bind_rows(dfES,dfTR)
df1<- bind_rows(df1,df2)
```

@@

@@

#Employment relation encuestado
#frq(df1$EMPREL)
#frq(df1$WORK)
df1$semp <- df1$EMPREL
df1$semp <- as.numeric(car::recode(df1$semp, "c(2,3,4)=1;c(1)=2;c(0,8,9)=NA"))
#frq(df1$semp)
df1$semp <- sjlabelled::set_labels(x = df1$semp,labels=c('Self-employed','Employee'))
#frq(df1$semp)

# superivison
#frq(df1$NEMPLOY) #supervision respondent
df1$supvis <- as.numeric(car::recode(df1$NEMPLOY,"c(9998,9999)=NA"))
#frq(df1$supvis)
# save the occupational variables for using them in stata-----------------------
df1$id <- NULL
df1 <- tibble::rowid_to_column(df1, "id")
haven::write_dta(data = df1[c("id","isco88r","semp","supvis",
                              "isco88spo","sempspo","supvisspo")],
                 path = here::here("input/data/proc/ocup_respondent-spouse.dta"))
```

```{r class-stata}
# Load the data created in stata------------------------------------------------
df_class <- haven::read_dta(file = here::here("input/data/proc/class_respondent-spouse.dta"))
df1 <- left_join(df1,df_class,by='id')

# Class R´s and Spouse----------------------------------------------------------
df1$class10 <- car::recode(df1$class10, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$class10 <- sjlabelled::set_labels(df1$class10, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                        labels = c("I", "II", "IIIa", "IIIb", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

sjlabelled::set_label(df1$class10) <- "Social Class - respondent"

df1$class10spo <- car::recode(df1$class10spo, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$class10spo <- sjlabelled::set_labels(df1$class10spo, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                        labels = c("I", "II", "IIIa", "IIIb", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

sjlabelled::set_label(df1$class10spo) <- "Social Class - spouse"

#if NA in class of respondent, replace with household

df1$class11 <- ifelse(is.na(df1$class10),df1$class10spo,df1$class10)
#frq(df1$class11)

# if NA in respondents it is assumed to be lower than spouse
# know, with the cases in we have information for respondent and spouse
# we replace the value by the highest one, just for those

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$class11d <- pmax(df1$class11, #social class respondent + spouse vs:
                       df1$class10spo,#social class spouse
                       na.rm = T) # if NA, do nothing

df1$class11d <- 
  sjlabelled::set_labels(df1$class11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                         labels =  c("I", "II", "IIIa", "IIIb", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
#frq(df1$class11d)
sjlabelled::set_label(df1$class11d) <- "Social Class"

@@

sjlabelled::set_label(df1$class3) <- "Social Class"
#frq(df1$class3)

@@

@@

```{r social-capital}
# Social capital measures
isei_ocupnames<- names(select(df1,starts_with("isei_")))

@@

```{r heterogeneity}
isei_ocupnames<- names(select(df1,busdriver:teacher))

@@

```{r homogeneity}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

```{r homogeneity-strong}
# EI homophily index
df1$ingroup_s<- NA 
df1 %>% select(class3,n_high_s,n_middle_s,n_low_s,ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high_s,df1$ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle_s,df1$ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low_s,df1$ingroup_s)
#frq(df1$ingroup_s)

@@

```{r homogeneity-weak}
# EI homophily index
df1$ingroup_w<- NA 
df1 %>% select(class3,n_high_w,n_middle_w,n_low_w,ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high_w,df1$ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle_w,df1$ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low_w,df1$ingroup_w)
# frq(df1$ingroup_w)

@@

```{r social distance}
df1$dis_busdriver <- df1$isei08 - df1$isei_busdriver
df1$dis_ceo <- df1$isei08 - df1$isei_ceo
df1$dis_cleaner <- df1$isei08 - df1$isei_cleaner
df1$dis_hairbarber <- df1$isei08 - df1$isei_hairbarber
df1$dis_hrmanager <- df1$isei08 - df1$isei_hrmanager
df1$dis_lawyer <- df1$isei08 - df1$isei_lawyer
df1$dis_mechanic <- df1$isei08 - df1$isei_mechanic
df1$dis_nurse <- df1$isei08 - df1$isei_nurse
df1$dis_police <- df1$isei08 - df1$isei_police
df1$dis_teacher <- df1$isei08 - df1$isei_teacher

```{r homogeneity-resp}
# EI homophily index
df1$ingroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Service Class (I+II)",df1$n_high,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup_res)
# frq(df1$ingroup_res)

@@

@@

```{r pca}
library(psych)
library(dplyr)
isei_ocupnames<- names(dplyr::select(df1,starts_with("isei_")))

@@

>>>>

# relig <- haven::read_dta(here::here("input/data/original/ZA7570_v2-1-0.dta"))
# religTR <- relig %>% filter(country==792)

df1 <- df1 %>% rename(topbot=TOPBOT)
dfES <- dfES %>% rename(topbot=TOPBOT)
#fix problem with variable names
dfTR <- dfTR %>% rename(SEX=sex,
                        DEGREE=degree,
                        EDUCYRS=educyrs,
                        AGE=age,
                        MAINSTAT=mainstat,
                        UNION=union,
                        WEIGHT=weight,
                        URBRURAL=urbrural,
                        WORK=work,
                        HOMPOP=hompop,
                        EMPREL=emprel,
                        NSUP=nsup,
                        WRKSUP=wrksup,
                        SPEMPREL=spemprel,
                        SPWRKSUP=spwrksup)  
dfES$c_alphan <- NULL
dfTR$c_alphan <- NULL
df2 <- bind_rows(dfES,dfTR)
df1<- bind_rows(df1,df2)
```

@@

# Index egalitarian attitudes
```{r}
df1 %>% select(redist,prefineq,perineq) %>% 
  psych::cor2()
df1 %>% select(redist,prefineq,perineq) %>% 
  psych::alpha()

df1$egal3<- rowMeans(df1[,c("redist","prefineq","perineq")],na.rm = T)
df1$egal3 <- ((df1$egal3-min(df1$egal3,na.rm = T))/(max(df1$egal3,na.rm = T)-min(df1$egal3,na.rm = T)))*100

sjlabelled::set_label(df1$egal3) <- "Redistributive attitudes (3 items)"
# hist.default(df1$egal3)

df1$egal2<- rowMeans(df1[,c("redist","prefineq")],na.rm = T)
df1$egal2 <- ((df1$egal2-min(df1$egal2,na.rm = T))/(max(df1$egal2,na.rm = T)-min(df1$egal2,na.rm = T)))*100
sjlabelled::set_label(df1$egal2) <- "Redistributive attitudes (2 items)"
# hist.default(df1$egal2)
```

@@

#Employment relation encuestado
frq(df1$EMPREL)
# frq(df1$WORK)
df1$semp <- df1$EMPREL
df1$semp <- as.numeric(car::recode(df1$semp, "c(2,3,4)=1;c(1)=2;c(0,8,9)=NA"))
frq(df1$semp)
df1$semp <- sjlabelled::set_labels(x = df1$semp,labels=c('Self-employed','Employee'))
#frq(df1$semp)

# superivison
frq(df1$NEMPLOY) #supervision respondent
frq(df1$NSUP)
df1$supvis <- as.numeric(car::recode(df1$NEMPLOY,"c(9998,9999)=NA"))

frq(df1$supvis)

# save the occupational variables for using them in stata-----------------------
df1$id <- NULL
df1 <- tibble::rowid_to_column(df1, "id")

# For comparing with the DIGCLASS

frq(df1$WRKSUP)
frq(df1$EMPREL)
frq(df1$NSUP)

df1 <-
  df1 %>%
  mutate(
    is_supervisor = ifelse(WRKSUP == 2, 0, as.numeric(car::recode(df1$WRKSUP,"c(8,9)=NA"))),
    self_employed = case_when(
      EMPREL %in% c(1,4) ~ 0,
      EMPREL %in% 2:3 ~ 1
    ),
    n_employees = ifelse(is_supervisor == 0, 0, as.numeric(car::recode(df1$NSUP,"c(9998,9999)=NA")))
  )

frq(df1$is_supervisor)
frq(df1$self_employed)
frq(df1$n_employees)

frq(df1$SPWORK)
frq(df1$SPEMPREL)

df1$NSUPSPO <- 1
frq(df1$NSUPSPO)

df1 <-
  df1 %>%
  mutate(
    is_supervisor_spo = ifelse(SPWORK == 2, 0, as.numeric(car::recode(df1$SPWORK,"c(8,9)=NA"))),
    self_employed_spo = case_when(
      SPEMPREL %in% c(1,4) ~ 0,
      SPEMPREL %in% 2:3 ~ 1
    ),
    n_employees_spo = ifelse(is_supervisor_spo == 0, 0, as.numeric(car::recode(df1$NSUPSPO,"c(9998,9999)=NA")))
  )

frq(df1$is_supervisor_spo)
frq(df1$self_employed_spo)
frq(df1$n_employees_spo)

df1 %>% select(isco88r, is_supervisor, self_employed,"semp", n_employees,"supvis")
haven::write_dta(data = df1[c("id","isco88r","semp","supvis",
                              "is_supervisor","self_employed","n_employees",
                              "isco88spo","sempspo","supvisspo",
                              "is_supervisor_spo","self_employed_spo","n_employees_spo")],
                 path = here::here("input/data/proc/ocup_respondent-spouse.dta"))
```

```{r class-from-stata}
# Load the data created in stata------------------------------------------------
df_class <- haven::read_dta(file = here::here("input/data/proc/class_respondent-spouse.dta"))
df1 <- left_join(df1,df_class,by='id')

# Class R´s and Spouse----------------------------------------------------------
frq(df1$class10)

# df1$class10 <- car::recode(df1$class10, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$class10f <- as_factor(df1$class10)
frq(df1$class10f)

df1$class10f <- factor(df1$class10f,
                       levels = c("higher controllers","lo controllers",
                                 "routine nonmanual",
                                 "sempl with emp","sempl without empl","selfempl farm",
                                 "manual supervisor","skilled manual","semi-unskilld manual","farm labor"),
                       labels = c("I", "II", "IIIab", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

df1$class10f <- as_labelled(df1$class10f)
sjlabelled::set_label(df1$class10f) <- "Social Class - respondent"
frq(df1$class10f)

df1$class10spof <- as_factor(df1$class10spo)
frq(df1$class10spof)

df1$class10spof <-factor(df1$class10spof,
                       levels = c("higher controllers","lo controllers",
                                 "routine nonmanual",
                                 "sempl with emp","sempl without empl","selfempl farm",
                                 "manual supervisor","skilled manual","semi-unskilld manual","farm labor"),
                       labels = c("I", "II", "IIIab", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

df1$class10spof <- as_labelled(df1$class10spof)
sjlabelled::set_label(df1$class10spof) <- "Social Class - spouse"

frq(df1$class10spof)
#if NA in class of respondent, replace with spouse

frq(df1$class10f)
frq(df1$class10spof)

df1$class11 <- ifelse(is.na(df1$class10f),df1$class10spof,df1$class10f)
frq(df1$class11)

# if NA in respondents it is assumed to be lower than spouse
# now, with the cases in we have information for respondent and spouse
# we replace the value by the highest one, just for those

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$class11d <- pmax(df1$class11, #social class respondent + spouse vs:
                     df1$class10spof,#social class spouse
                     na.rm = T) # if NA, do nothing
frq(df1$class11d)
df1$class11d <- 
  sjlabelled::set_labels(df1$class11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                         labels =  c("I", "II", "IIIab", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
frq(df1$class11d)
sjlabelled::set_label(df1$class11d) <- "Social Class"

@@

sjlabelled::set_label(df1$class3) <- "Social Class"
frq(df1$class3)

@@

```{r class-from-stata2}
# Load the data created in stata------------------------------------------------
# Class R´s and Spouse----------------------------------------------------------
df1$digclass10 <- car::recode(df1$digclass10, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$digclass10 <- sjlabelled::set_labels(df1$digclass10, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                        labels = c("I", "II", "IIIa", "IIIb", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

sjlabelled::set_label(df1$digclass10) <- "Social Class - respondent"

df1$digclass10spo <- car::recode(df1$digclass10spo, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$digclass10spo <- sjlabelled::set_labels(df1$digclass10spo, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                        labels = c("I", "II", "IIIa", "IIIb", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

sjlabelled::set_label(df1$digclass10spo) <- "Social Class - spouse"

#if NA in class of respondent, replace with spouse

df1$digclass11 <- ifelse(is.na(df1$digclass10),df1$digclass10spo,df1$digclass10)
#frq(df1$class11)

# if NA in respondents it is assumed to be lower than spouse
# now, with the cases in we have information for respondent and spouse
# we replace the value by the highest one, just for those

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$digclass11d <- pmax(df1$digclass11, #social class respondent + spouse vs:
                       df1$digclass10spo,#social class spouse
                       na.rm = T) # if NA, do nothing

df1$digclass11d <- 
  sjlabelled::set_labels(df1$digclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                         labels =  c("I", "II", "IIIa", "IIIb", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
frq(df1$digclass11d)
sjlabelled::set_label(df1$digclass11d) <- "Social Class"

# Class Scheme 3 classes
df1$digclass3 <- car::recode(df1$digclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3) <- "Social Class"
frq(df1$digclass3)

# Class Scheme 5 classes
df1$digclass5 <- car::recode(
  df1$digclass11d,
  as.factor = T,
  levels = c(
    'Service class (I+II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I+II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$digclass5) <- "Social Class"
frq(df1$digclass5)
# Class Scheme 6 classes
df1$digclass6 <- car::recode(
  df1$digclass11d,
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

# Class only R's --------------------------------------------------------------
df1$digclass10

# Class Scheme 3 classes
df1$digclass3res <- car::recode(as.numeric(df1$digclass10),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3res) <- "Social Class (R's)"
# frq(df1$class3res)

# Class Scheme 5 classes
df1$digclass5res <- car::recode(
  as.numeric(df1$digclass10),
  as.factor = T,
  levels = c(
    'Service class (I+II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I+II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$digclass5res) <- "Social Class (R's)"
# frq(df1$class5res)

# Class Scheme 6 classes
df1$digclass6res <- car::recode(
  as.numeric(df1$digclass10),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$digclass6res) <- "Social Class (R's)"
frq(df1$digclass6res)
# Class only Spouse --------------------------------------------------------------
df1$digclass10spo

# Class Scheme 3 classes
df1$digclass3spo <- car::recode(as.numeric(df1$digclass10spo),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3spo) <- "Social Class (Spouse)"
frq(df1$digclass3spo)

# Class Scheme 5 classes
df1$digclass5spo <- car::recode(
  as.numeric(df1$digclass10spo),
  as.factor = T,
  levels = c(
    'Service class (I+II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I+II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$digclass5spo) <- "Social Class (Spouse)"
# frq(df1$class5res)

# Class Scheme 6 classes
df1$digclass6spo <- car::recode(
  as.numeric(df1$digclass10spo),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)
sjlabelled::set_label(df1$digclass6spo) <- "Social Class (Spouse)"
```

# Class using R

```{r digclass}
# install.packages("devtools")
# devtools::install_git("https://code.europa.eu/digclass/digclass.git")
library("DIGCLASS")
library(dplyr)

df1 <- 
df1 %>% 
mutate(egp11 = isco88_to_egp(isco88r, self_employed, n_employees, label = F,to_factor = T),
       egp11spo = isco88_to_egp(isco88spo, self_employed_spo, n_employees_spo, label = F,to_factor = T))

frq(df1$egp11)
df1$egp11 <- as_numeric(df1$egp11)
frq(df1$egp11)

frq(df1$egp11spo)
df1$egp11spo <- as_numeric(df1$egp11spo)
frq(df1$egp11spo)

#if NA in class of respondent, replace with spouse
df1$dclass11 <- ifelse(is.na(df1$egp11),df1$egp11spo,df1$egp11)

# if NA in respondents it is assumed to be lower than spouse
# now, with the cases in we have information for respondent and spouse
# we replace the value by the highest one, just for those

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$dclass11d <- pmax(df1$egp11, #social class respondent + spouse vs:
                     df1$egp11spo,#social class spouse
                     na.rm = T) # if NA, do nothing
frq(df1$dclass11d)
df1$dclass11d <- as_numeric(df1$dclass11d)
frq(df1$dclass11d)

df1$dclass11d <- 
  sjlabelled::set_labels(df1$dclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,11),
                         labels =  c("I", "II", "IIIa","IIIb", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
frq(df1$dclass11d)
sjlabelled::set_label(df1$dclass11d) <- "Social Class"

# Class Scheme 3 classes
df1$dclass3 <- car::recode(df1$dclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$dclass3) <- "Social Class"
frq(df1$dclass3)
frq(df1$class3)
```

@@

```{r social-capital-measures}
# Social capital measures
isei_ocupnames<- names(select(df1,starts_with("isei_")))

@@

```{r heterogeneity, eval=FALSE, include=FALSE}
isei_ocupnames<- names(select(df1,busdriver:teacher))

@@

```{r homogeneity-v01}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

```{r homogeneity-strong-v01}
# EI homophily index
df1$ingroup_s<- NA 
df1 %>% select(class3,n_high_s,n_middle_s,n_low_s,ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high_s,df1$ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle_s,df1$ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low_s,df1$ingroup_s)
#frq(df1$ingroup_s)

@@

```{r homogeneity-weak-v01}
# EI homophily index
df1$ingroup_w<- NA 
df1 %>% select(class3,n_high_w,n_middle_w,n_low_w,ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high_w,df1$ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle_w,df1$ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low_w,df1$ingroup_w)
# frq(df1$ingroup_w)

@@

```{r homogeneity-resp-v01}
# EI homophily index
df1$ingroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Service Class (I+II)",df1$n_high,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup_res)
# frq(df1$ingroup_res)

@@

```{r homogeneity-v02}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$dclass3=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex2<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex2) <- "EI homophily index"
#frq(df1$eindex)
hist.default(df1$eindex)
df1$eindex_gc <- group_center(x = df1$eindex,grp = df1$country)
hist.default(df1$eindex_gc)

df1 %>% select(dclass3,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass2 <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass2) <- "Class homogeneity"
df1$homclass2_gc <- group_center(x = df1$homclass2,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

@@

```{r pca, eval=FALSE, include=FALSE}
library(psych)
library(dplyr)
isei_ocupnames<- names(dplyr::select(df1,starts_with("isei_")))

@@

```{r social distance, eval=FALSE, include=FALSE}
df1$dis_busdriver <- df1$isei08 - df1$isei_busdriver
df1$dis_ceo <- df1$isei08 - df1$isei_ceo
df1$dis_cleaner <- df1$isei08 - df1$isei_cleaner
df1$dis_hairbarber <- df1$isei08 - df1$isei_hairbarber
df1$dis_hrmanager <- df1$isei08 - df1$isei_hrmanager
df1$dis_lawyer <- df1$isei08 - df1$isei_lawyer
df1$dis_mechanic <- df1$isei08 - df1$isei_mechanic
df1$dis_nurse <- df1$isei08 - df1$isei_nurse
df1$dis_police <- df1$isei08 - df1$isei_police
df1$dis_teacher <- df1$isei08 - df1$isei_teacher
